Case-based Arabic Morphological Analysis

### 2010-12-06 tex/rp.tex
method

A \RL{.hady_t} is a narration related to the prophet Mohammad
through a \RL{sanad} or a sequence of narrators. 
Figure~\ref{f:exhadith} shows an example \noArRL{.hady_t} in Arabic with its 
transliteration and translation. 
The authentication of a \noArRL{.hady_t} highly depends on the credibility
of each of the narrators as reported in separate biography 
books. 
In this paper we consider the problem of automatically segmenting
a \noArRL{.hady_t} book into narrations, then segmenting each 
narration into
its content or \RL{matn} and its \noArRL{sanad}.
We also partition the sanad accurately into the 
separate narrators so that we can later look each one of them 
up in the biography books. 
For example, the boxes in Figure~\ref{f:exhadith} are proper names 
and once
connected toghether they form complex names of narrators. 
Narrator $n_1$ has the first name \noTrRL{qtybT} and his father 
name is \noTrRL{s`yd} as the word in between 
\noTrRL{bn} (son of) indicates.

@@

The directed acyclic graph (DAG) 
in Figure~\ref{f:narrators} shows a partial order relation between
narrators that was automatically extracted from hadith text 
documents.
Given a set of chains of narrators similar to 
Figure~\ref{f:exhadith} we formed the DAG using graph algorithms 
that merged equal names in the chains. 
This partial order graph is instrumental to automatically
locate narrators in biography documents and automatically
segment biography documents.

@@

Text mining concerns the partitioning of text segments into classes,
the clustering of text segments,
the extraction of concepts and entities from text,
and the discovery and modeling of relations between text entities 
and classes.
Text mining techniques perform NLP to analyze text and perform 
IE to extract information into data structures.
Fundamental to IE are the name entity (NE) and the name entity 
relation (NER) extraction techniques which capture features from 
sentences and phrases.
A key challenge to NE and NER is that sentence structures and 
words are often ambiguous in natural languages.
While research to address ambiguity in Latin languages is still 
lagging behind~\cite{Red08},
not much has been done to address NE and NER in the context of 
the Arabic language.

@@

The difficulty of applying text mining techniques to Arabic text 
documents lays in the absence of automated tools that understand 
the unique features of the Arabic language.
We will make use of an example to illustrate the process.
Given an Arabic text document we want a tool that can segment 
and isolate all names, dates, time of the day occurrences,
tool names, and locations as entities.
Then we want another tool to identify relations amongst 
these entities such as identities and orders.
An order between two entities can be simply 
their order of appearance in the document, or their
numerical 
 We want another tool to look at the verbs and actions in the sentences and try to build relations between the identified entities based on these verbs.
 If we were successful to do that,
 we now want to order all entities associated with dates based on a chronological order.

>>>>

A \RL{.hady_t} is a narration related to the prophet Mohammad
through a \RL{sanad} or a sequence of narrators. 
Figure~\ref{f:exhadith} shows an example \noArRL{.hady_t} in Arabic with its 
transliteration and translation. 
The authentication of a \noArRL{.hady_t} highly depends on the credibility
of each of the narrators as reported in separate biography 
books. 
We consider the problem of automatically segmenting
a \noArRL{.hady_t} book into narrations, then segmenting each 
narration into
its content or \RL{matn} and its \noArRL{sanad} or the
chain of narrators.
We also partition the sanad accurately into the 
separate narrators so that we can later look each one of them 
up in the biography books. 
For example, the boxes in Figure~\ref{f:exhadith} are proper names 
and once
connected toghether they form complex names of narrators. 
Narrator $n_1$ has the first name \noTrRL{qtybT} and his father 
name is \noTrRL{s`yd} as the word in between 
\noTrRL{bn} (son of) indicates.

@@

The directed acyclic graph (DAG) 
in Figure~\ref{f:narrators} shows a partial order relation (POR) between
narrators.
We automatically extracted the POR from hadith text 
documents.
The nodes in boxes are the \RL{matn} of the hadith, 
and the other nodes are the narrators.
Given a set of chains of narrators similar to 
Figure~\ref{f:exhadith} we formed the DAG using graph algorithms 
that merged equal names in the chains. 
This partial order graph is instrumental to automate
localizing narrators in biography documents and
segmenting biography documents.

Given that a biography will discuss a narrator and mention
his professors and students,
we can segment the biographies with a graph coloring algorithm 
that traverses the text and colors the POR whenever
a name is found. 
We modify the color when we move in the graph 
far away from the last name we colored.

Once the biographies are analyzed, one can annotate
the POR with qualifiers of the narrators that reflect
their authenticity. 
One can also annotate the POR with the locations and 
the time they lived in. 
Then One can compute a time and location overlap
check to discover inconsistent narrations.
We can also perform several interesting checks using 
this POR on its own such as checking for the effect of
one narrator on the hadith literature.

We think that the ability to extract such orders
between entities in documents can

@@

Text mining concerns the partitioning of text segments into classes,
the clustering of text segments,
the extraction of concepts and entities from text,
and the discovery and modeling of relations between text entities 
and classes.
Text mining techniques perform NLP to analyze text and perform 
IE to extract information into data structures.
Fundamental to IE are the named entity (NE) and the named entity 
relation (NER) extraction techniques which capture features from 
sentences and phrases.
A key challenge to NE and NER is that sentence structures and 
words are often ambiguous in natural languages.
%While research to address ambiguity in Latin languages is still 
%lagging behind~\cite{Red08},
%not much has been done to address NE and NER in the context of 
%the Arabic language.

@@

The difficulty of applying text mining techniques to Arabic text 
documents lays in the absence of automated tools that understand 
the unique features of the Arabic language.
%We will make use of an example to illustrate the process.
Given an Arabic text document we want a tool that can segment 
and isolate all names, dates, time of the day occurrences,
tool names, and locations as entities.
Then we want another tool to identify relations amongst 
these entities such as identities and orders.
An order between two entities can be simply 
their order of appearance in the document, or their
numerical 
 We want another tool to look at the verbs and actions in the sentences and try to build relations between the identified entities based on these verbs.
 If we were successful to do that,
 we now want to order all entities associated with dates based on a chronological order.
